### 2014-12-08 Rebuttal.md
added Rebuttal file for tracking changes due to reviews

# Rebuttal

## Reviewer 1.
- Applicability: Comparison to standard workflow is not possible. The classic epidemiological workflow is strictly hypothesis driven and therefore requires a small set of 2 - 3 variables, which are correlated with a target variable. IVA methods amplify *hypothesis generation* and therefore aim to provide insight into unexpected relationships in the data.