### 2014-12-08 Rebuttal.md
minor revisions at draft and added comments to the Rebuttal

Rebuttal
========

This rebuttal is written in markdown.

Reviewer 1
----------

- Figure 1 was completely removed since it does not supports the train of thought as we intended and due to lack of space
- Applicability: Comparison to standard workflow is not possible. The classic epidemiological workflow is strictly hypothesis driven and therefore requires a small set of 2 - 3 variables, which are correlated with a target variable. IVA methods amplify *hypothesis generation* and therefore aim to provide insight into unexpected relationships in the data.
- Exploration process is unclear
	- at the beginning of Section 5 - Experiments and Preliminary Results, all subjects