### 2014-12-16 draft.tex
removed comments from the LaTeX source

\documentclass[a4paper,twoside]{style/article}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{calc}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{multicol}
\usepackage{pslatex}
\usepackage{apalike}
\usepackage{style/SCITEPRESS}
\usepackage[small]{caption}

\subfigtopskip=0pt
\subfigcapskip=0pt
\subfigbottomskip=0pt

% My Packages and Commands ---
\usepackage[normalem]{ulem}
\usepackage{xcolor}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[utf8]{inputenc}
\usepackage{url}
\newcommand{\rem}[1]{\textcolor{red}{\sout{#1}}}
\newcommand{\add}[1]{\textcolor{blue}{\uline{#1}}}
\newcommand{\com}[1]{\textcolor{orange}{\uline{#1}}}
% /My Packages and Commands --

\begin{document}

\title{Interactive Visual Analysis of Lumbar Back Pain  \subtitle{What the Lumbar Spine Tells About Your Life} }

\author{\authorname{Paul Klemm\sup{1}, Sylvia Glaßer\sup{1}, Kai Lawonn\sup{1}, Marko Rak\sup{1}, Henry Völzke\sup{2}, Katrin Hegenscheid\sup{3} and Bernhard Preim\sup{1}}
\affiliation{\sup{1} Department of Simulation and Graphics, University of Magdeburg, Germany}
\affiliation{\sup{2} Institute for Community Medicine, University of Greifswald, Germany}
\affiliation{\sup{3} Institute for Diagnostic Radiology and Neuro-Radiology, University of Greifswald, Germany}
\email{klemm@isg.cs.uni-magdeburg.de}
}

\keywords{Epidemiology, Interactive Visual Analysis, Classification, Multi-Modal Data}

\abstract{ 
Epidemiology aims to provide insight into disease causations.
%%
Hence, subject groups (cohorts) are analyzed to correlate the subjects' varying lifestyles, their medical properties and diseases.
%%
Recently, these cohort studies comprise medical image data.
%%
We assess potential relations between image-derived variables of the lumbar spine with lower back pain in a cross-sectional study.
%%
Therefore, an Interactive Visual Analysis (IVA) framework was created and tested with 2,540 segmented lumbar spine data sets.
%%
The segmentation results are evaluated and quantified by employing shape-describing variables, such as spine canal curvature and torsion.
%%
We analyze mutual dependencies among shape-describing variables and non-image variables, e.g., pain indicators.
%%
Therefore, we automatically train a decision tree classifier for each non-image variable.
%%
We provide an IVA technique to compare classifiers with a \emph{decision tree quality plot}.
%%
As a first result, we conclude that image-based variables are only sufficient to describe lifestyle factors within the data.
%%
A correlation between lumbar spine shape and lower back pain could not be found with the automatically trained classifiers.
%%
However, the presented approach is a valuable extension for the IVA of epidemiological data.
%%
Hence, relations between non-image variables were successfully detected and described.
}

\onecolumn \maketitle \normalsize \vfill

\section{\uppercase{Introduction}}
\label{sec:Introduction}

\noindent Epidemiology is the study of dissemination, causes and results of health-related states and events.
%%
Large population studies, such as the \emph{Study of Health in Pomerania} (SHIP) \cite{SHIP}, gather as much information as possible about participants to be assessed towards different diseases.
%%
This information is used to determine risk factors for diseases, inform people about healthier lifestyles or to support the diagnosis of widespread diseases.
%%
Cohort studies are an instrument of epidemiological research.
%%
We analyze back pain, one of the most frequent complaints in the Western civilization.
%%
Although the shape and constitution of the spine, especially the lumbar spine, plays an important role for back pain, an automatic classification approach for characterization of back pain based on lumbar spine attributes is still missing.
%%

We present an analysis of the image-derived data from a cohort study lumbar spine dataset.
%%
We extract possible associations between spine shape and back pain characteristics.
%%
For this purpose, we combine classification algorithms with data visualization techniques.
%%
Then, Interactive Visual Analysis (IVA) highlights mutual dependencies between image-derived data and back pain-related variables.
%%
We focus on highlighting new correlations and trigger \emph{hypotheses generation}, rather than statistically validate complex epidemiological correlations.
%%
Our contributions are:
\begin{itemize}
\item An IVA workflow for back pain analysis based on image-derived variables of 2,240 subjects,
%%
\item The identification of lumbar spine shape properties potentially associated with back pain,
%%
\item The detection of associations between image-based, socio-demographic and medical variables for \emph{hypotheses generation},
%%
\item The identification of the most important variables via classification methods using a novel \emph{decision tree quality plot}.
%%
\end{itemize}
%% ---------------------------------------------
\section{\uppercase{Epidemiological Background}}
\label{sec:EpidemiologicalBackground}
%%
\noindent Epidemiological reasoning relies on a strict statistically-driven workflow \cite{Fletcher}:
%%
(1) physicians formulate hypotheses based on observations made in their clinical practice; (2) epidemiologists compile a list of variables depicting a hypothesis to prove it; (3) statistical methods, such as regression analysis, assess the correlation of selected variables with the investigated condition.
%%
Mutually dependent variables make this analysis challenging.
%%
For example, the incidence rate of many diseases, such as different cancer types increases with age \cite{Fletcher}.
%%------------------------------
\paragraph{Challenges of Epidemiological Data Analysis.}
%%
We focus on epidemiological large scale cohort study data.
%%
The collected data can be analyzed regarding many diseases and conditions.
%%
Epidemiological data are heterogenous and incomplete.
%%
For example, data about a disease treatment only affects subjects suffering from this condition.
%%
Statistical analysis has to take missing data into account.
%%
Epidemiological data acquisition includes a variety of techniques, such as medical examinations, self-reported questionnaires or genetic examinations.
%%
This yields a heterogenous information space.
%%
Information reduction techniques are necessary to compare these data.
%%
For example, continuous data, such as age, is often discretized into quantile bins for visualizing age-dependent disease frequencies.

Modern cohort studies often comprise medical image data.
%%
The analysis of these data is challenging.
%%
Individual diagnosis or manual segmentation of each body structure by radiologists is tedious, costly and comprises little reproducibility.
%%
Segmentation algorithms are not generally available and need to be carefully adapted for each body structure.
%%--------------------------------
\\\\
\noindent \textbf{Lower Back Pain.}
%%
Lower back pain is one of the most frequent complaints in the Western civilization \cite{Hoy2010}.
%%
The exact causes as well as particular vulnerable risk groups are unknown.
%%
Epidemiologists want to describe the relation between aging and degenerative process of the spine \cite{Szpalski2005}.
%%
They have to analyze the lumbar spine shape as well as lifestyle factors.
%% -------------------------------
\section{\uppercase{Related Work}}
\label{sec:RelatedWork}
%%
\noindent \textbf{Visual Analysis of Medical Image and Non-Image Data.}
%%
In a previous work, we proposed an epidemiological IVA workflow, consisting of an iterative sequence of \emph{group selection}, \emph{variable selection} and \emph{visualization} \cite{Klemm2014VIS}.
%%
\emph{Hypotheses generation} was amplified by concurrently analyzing heterogenous variables at once using correlation measures.
%%
\cite{Turkay} derived descriptive data metrics from image data.
%%
Their proposed \emph{deviation plot} shows distribution-specific measures of a variable, such as skewness or inter-quantile range, making variables comparable.
%%
This approach aimed to trigger \emph{hypotheses generation} by outlining tendencies between these variables.
%%
A survey on image-centric cohort studies and strategies to analyze the resulting data is given in \cite{Preim2015}.
%%
In a prior work, we analyzed the lumbar spine variability of 490 subjects \cite{Klemm2013VMV}.
%%
We incorporated hierarchical agglomerative clustering to derive shape groups, yielding average shape groups and outliers.
%%
As opposed to clustering the spine shape, we analyze the discriminative power of the resulting data towards back pain and other non-image variables in this work.
%%
\\\\
\noindent \textbf{Visual Analysis of Heterogenous Non-Image Data.}
%%
\cite{Zhang} analyzed subject groups in a web-based linked view system.
%%
The resulting decision rules aim to categorize new subjects as they are added to the data.
%%
They defined a cohort as variable-divided subject group, differing from the epidemiological understanding of the term.
%%
The described method lacks details about handling of missing data, the definition of similarity or the choice of the statistical measures.
%%
Generalized Pairs Plots (\texttt{GPLOMs}) visualize heterogenous variables in a plot matrix \cite{GPLOMS}. %\cite{GPLOMS,Francois}.
%%
The plot depends on the type of variables, which are visualized pairwise.
%%
\texttt{GPLOMs} provide an overview, but take up much screen space and are therefore only suitable for a few variables at once (see example in Fig.~\ref{fig:image-parameter-range}).
%%----------------------------
\\\\
\noindent \textbf{Decision Rule-Driven Analysis of Medical Data.}
%%
Closest to ours is the work of \cite{Glasser2013} and \cite{Niemann2014}.
%%
Both used decision trees for their analyses.
%%
Decision trees are easily readable and are frequently used for classifying medical data.
%%
\cite{Glasser2013} used variables derived from DCE-MRI data capturing the perfusion in tumorous tissue.
%%
To classify breast tumors, they trained a decision-tree classifier, concluding that the extracted kinetic and morphological variables alone are not sufficient for tumor type classification.
%%
\cite{Niemann2014} assessed risk factors for hepatic steatosis (fatty liver disease) using decision trees.
%%
Their interactive data mining tool can analyze association rules and highlights relations.
%%
We combine both ideas by validating the significance of image-derived variables.
%%
\\\\
Unique in our work is the combination of classification techniques with an IVA approach by observing interesting variable relations in the context of image-derived variables using multiple decision trees for one visualization.
%%
We abstract the decision tree results similar to \cite{Turkay}, making them comparable in an overview visualization.
%%
\section{\uppercase{The Lumbar Spine Data Set}}
\label{sec:MaterialsAndMethod}
%%
\noindent Our work is based on a data set compiled by Epidemiologists with a wide range of variables possibly correlating with lumbar back pain, comprising non-image and image-derived data for 6,753 subjects.
%%
\\\\
\noindent \textbf{Non-Image Data}.
%%
The non-image variables range from somatometric variables describing body measures to medical examinations, such as laboratory tests as well as lifestyle factors, e.g., sporting activity or nutrition.
%%
The data set comprises 134 variables: %(and 9 additional image-derived variables described in the next subsection):
\begin{itemize}
	\item 21 metric variables, describing somatometric variables and markers retrieved from blood analysis.
	%%
	\item 113 categorical variables divided into 43 dichotomous (binary) variables, mostly indicating the presence of a disease, e.g., \emph{pancreatitis} or \emph{high blood pressure}, and 70 variables with more than two levels, indicating pain levels, nutrition and social factors, such as marital or retirement status.
	%%
\end{itemize}
%%
All categorical variables are converted into binary \emph{dummy variables}, indicating the presence or absence of a categorical variable manifestation.
%%
For example, a pain indicator variable ranging from \emph{1 - no pain} to \emph{4 - large pain} is subdivided into four dichotomous variables to determine, which manifestation can be described best using the image-based variables.
%%
\\\\
\noindent \textbf{Image-Derived Data.}
\begin{figure}[!t]
  %\vspace{-0.2cm}
  \centering
  \includegraphics[width=0.475\textwidth]{figures/centerline}
  \caption{
	(a) Finite element model (FEM) of the lumbar spine (left), capturing the L1-L5 vertebrae and the lumbar spine canal (right).
	%%
	The purple dashed line describes the lumbar spine canal centerline with 92 points.
	%%
	(b) We extract the weighted sum of \emph{curvature} and \emph{torsion} for all 92 points (white dashes) and the \emph{curvature angle} ($\alpha$) for each projection axis to assess their information gain. %is calculated using the top, bottom and middle point.
	%%
	}
  \label{fig:centerline}
\end{figure}
Magnetic Resonance Imaging (MRI) scans were obtained for each subject \cite{Hegenscheid2013}.
%%
A hierarchical finite element method was used to detect the lumbar spine in the MRI data \cite{Rak2013}.
%%
The tetrahedron-based finite element model (Fig.~\ref{fig:centerline}~a) captures information about the lumbar spine canal shape and the position of the L1-L5 vertebrae.
%%
The detection fails for several subjects due to imaging artifacts or strongly deformed spines, yielding models for $2,540$ out of $6,753$ subjects.
%%
The detection model depicts the vertebrae positions, spine canal curvature, but lacks detailed information about their volume.
%%
In \cite{Klemm2013VMV}, we extracted a centerline representation of the lumbar spine canal from the detection model (Fig.~\ref{fig:centerline}~a).
%%
Using the Frenet frame, we calculated the following metrics from the model (Fig.~\ref{fig:centerline}~b): %\cite{Frenet}:
%%
\begin{itemize}
	\item \emph{Mean Curvature} is defined as the average curvature of all points describing the centerline: $\sum_{i=1}^I \frac{\textit{curvature}_i}{I}$. We refer to the mean curvature as \emph{curvature}.
%%
	\item \emph{Mean Torsion} (deviation of a curve from its current course) is defined as the average torsion of all points describing the centerline: $\sum_{i=1}^I \frac{\textit{torsion}_i}{I}$. We refer to the mean torsion as \emph{torsion}.
%%
	\item \emph{Curvature angle $\alpha$} is the angle defined by the middle point of the spine canal centerline as \emph{vertex} and the line connecting middle point and top/bottom point as \emph{sides}.
%%
\end{itemize}
%%
These metrics are also extracted in the sagittal, coronal and transversal projection of the model, yielding 9 image-derived variables.
%%
In the next section, we present the experiments we conducted to assess the influence of the lumbar spine shape to lower back pain.
%%
\section{\uppercase{Experiments and Preliminary Results}}
\label{sec:Experiments}
%%
\noindent In this section, the image-derived variables are analyzed towards the dichotomous \emph{back pain} indicator using a \texttt{GPLOM} and all non-image variables using heterogenous correlations.
%%
Spine shape is influenced by several somatometric variables.
%%
Larger people also have a longer spine with a straighter shape.
%%
High \emph{body weight} increases the spine load, resulting in a bent shape.
%%
To assess their influence, we take them into account when spine \emph{curvature} and \emph{torsion} are correlated with non-image variables.
%%
Discretizing metric variables using quartiles avoids outlier groups.
%%
\\\\
\noindent \textbf{GPLOM Analysis.}
As first experiment we correlated the shape variable with the dichotomous back pain indicator using a \texttt{GPLOM} (Fig.~\ref{fig:image-parameter-range}).
%%
The metric image-derived variables are pairwise visualized using scatter plots on the left side of the matrix diagonal.
%%
The combination of the image variables with back pain is visualized as histogram in the last row and as box plot in the last column.
%%
The projections to the transversal planes attract attention as they have many outliers.
%%
We conclude that \emph{curvature} is not as reliable on the transversal plane as it is on the other planes, which was also confirmed through a principal component analysis (see supplementary material at \url{ivapp15.dnsalias.com}).
%%
The \texttt{GPLOM} shows similar distributions of subjects with or without back pain with respect to the shape variables in all sub-plots.
%%--------------------------------
\begin{figure*}[htb]
  \centering
  \includegraphics[width=1.0\textwidth]{figures/image-parameter-range}
  \caption{
%%
A \texttt{GPLOM} of all image-derived variables colored by presence (red) or absence (turquoise) of back pain.
%%
Pairwise combinations of image-derived variables are visualized via scatter plots on the left of the matrix diagonal.
%%
Their correlation with back pain is denoted to the right of the matrix diagonal.
%%
The box plots (right) and histograms (bottom) display the distribution of each image variable encoded with back pain.
%%
No correlations with back pain can be identified in this plot.
}
  \label{fig:image-parameter-range}
\end{figure*}
%%
\\\\
\noindent \textbf{Heterogenous Correlations.}
%%
We then expanded our focus on correlations of image-derived variables with all other non-image variables.
%%
Different correlation metrics depending on the type of the individual variables are used to derive correlations between all variables in the data set.
%%
The method uses the \emph{Pearson product-moment} for two continuous variables, \emph{polyserial} correlation for one continuous and one categorical variable and \emph{polychoric} correlations for two categorical variables.
%%
All values are scaled between \texttt{0 - no correlation} and \texttt{1 - perfect correlation}.
%%
Some variables are too sparse for calculating correlations, for example \emph{treatment of diabetes}, or \emph{medication against high blood pressure} are omitted, since they are not statistically resilient.
%%
We display the resulting \emph{contingency matrix} using a heat map, encoding correlation values with color brightness with white for $0$ and dark blue for $1$.
%%
We calculated the contingency matrix for all size groups and searched for correlations between image- and non-image variables.
%%
The resulting contingency matrices show no strong correlation with image variables (see experiments page at \url{ivapp15.dnsalias.com}).
%%
Only weak correlations could be found for \emph{mean curvature} with \emph{gender} (0.42), \emph{body size} (0.39) and \emph{number of born children} (0.29).
%%
One surprising result was the small correlation of \emph{torsion} with \emph{Parkinson's disease} (0.24).
%%
Other than that, \emph{torsion} correlated with almost no variables (values between 0 and 0.05).
%%

These observations brought us to the decision to incorporate classification techniques to assess the influence of the image-derived variables.
%%
\section{\uppercase{Interactive Decision Tree Quality Plot}}
\label{sec:DecisionTrees}
%%
\noindent As described before, correlation coefficients fail to infer back pain status based on lumbar spine canal \emph{curvature} and \emph{torsion}.
%%
We rely on predictive classification to obtain a complex rule set on how combinations of the image-variables explain non-image variables.
%%
Decision trees are used to create predictive models.
%%
These models are built w.r.t. all input variables and capture more complex relationships than correlation coefficients.
%%
Leafs of a decision tree represent class labels, branches represent variable conjunctions leading to the class labels.
%%
Decision trees are easy to understand and to read.
%%
Too many branches impose overfitting to the data \cite{DecisionTree}.
%%
\\\\
\noindent \textbf{Creating Decision Trees.}
We use the C4.5 algorithm, which builds decision trees based on information entropy.
%%
Categorical attributes with more levels are biased with more information gain in a decision tree \cite{deng2011bias}.
%%
Creating a dummy variable by converting each manifestation of a categorical variable into a dichotomous variable bypasses this problem.
%%
In the following analysis, we strongly focus on the complexity of decision trees and the classification accuracy.
%%

We have to create a decision tree for every non-image variable to analyze which one can be explained by image-derived variables.
%%
Since we have 134 non-image variables, the calculation yields the same amount of trees.
%%
Further subdivision, e.g., by quantiles of \emph{body mass index}, increases the number to 402 trees.
%%
We have to abstract the results of the classification to keep the mental effort of interpreting the data low.
%%------------
\\\\
\noindent \textbf{Decision Tree Quality Plot.}
%%
We follow the Visual Analytics mantra of \emph{analyzing first}, \emph{show the important} and \emph{analyze further} \cite{Keim}.
%%
A first analysis step was performed by applying the classification algorithm to the data.
%%
The optimal classification uses a few rules to precisely describe the target variable.
%%
Therefore, we are interested in \emph{small trees} with a \emph{low classification error}.
%%
The two measures form the axes for a scatter plot of the classification results.
%%
This \emph{decision tree quality plot} is our central element for the interactive analysis of decision trees.
\\\\
\noindent \textbf{The Error Term.}
Calculating the mean classification error is imprecise for non-uniform distributions.
%%
For example, if a variable indicating a disease is negative for 90\% of the subjects and the classifier simply assigns all subjects to \emph{not ill}, we get a mean error of 10\%, even though it is very imprecise.
%%
Based on this we use a summary error based on the weighted mean, which incorporates the discriminative power of each manifestation and is denoted as follows:
%%
\begin{equation}
\textit{totalError} = 1 - \sum_{m=1}^M \frac{\textit{correctlyClassified}_{m}}{M\cdot N_m}
\end{equation}
%%
$M$ represents the set of manifestations of each variable.
%%
$N_m$ denotes the number of subjects in manifestation $m$.
%%
The error represents the share of incorrect classifications and denotes perfect classification with 0 and always wrong with 1. %; and random selection with 0.5.
%%
We display results below $0.5$ in the visualization, a value below $0.25$ represents a good classification.
%%
It allows for comparability of error rates between variables with different manifestation count.
\\\\
\textbf{Attribute Mapping.}
%%
The scatter plot axes are defined by \emph{tree size} and the previously described \emph{error metric}.
%%
This allows us to visualize a multitude of classification results in one plot.
%%
Classification and comparison of variables for subject groups (e.g., male and female subjects) in one plot can be achieved by color coding group affiliation on the data points.
%%
Many variables are sparse, such as \emph{medication of diabetes} or \emph{reason of early retirement}.
%%
The classification algorithm may produce higher accuracy for variables with less subjects due to the small sample size, making these results less reliable.
%%
Therefore, we provide a way to adjust the minimal number of subjects for each variable using a slider.
%%
The initial value is empirically set to $100$, marking a good tradeoff between sparse variables and statistical informative value.
%%
Furthermore, we map the number of subjects associated with a variable to the diameter in the decision tree quality plot.
%%
This allows instant reliability assessment of the result.
%%
We apply a square root scale for the tree size axis to highlight decision trees with few decision rules.
%%
Outlier results with large decision trees would otherwise distort the resulting plot.
%%------------
\\\\
\noindent \textbf{Decision Tree Quality Plot Interaction.}
%%
The visualization provides a good overview of the classification results.
%%
\emph{Details-on-demand} are displayed by clicking on an entry in the visualization, which then displays the corresponding decision tree in detail.
%%
This allows to sequentially analyze the classifications.
%%
We provide controls for adjusting the maximum classification error and minimum subject count for a variable.
%%
This gives the user control to abstract or refine the displayed information.
%%
The subject subdivision is controlled by the selection of variables, such as \emph{gender} or \emph{employment status}.
%%
Metric variables, such as \emph{body size}, are discretized using their quantiles.
%%
This allows to assess the influence of a variable to the classification process.
%%------------
\\\\
\noindent \textbf{Implementation.}
All analyses are available at \url{ivapp15.dnsalias.com} and are carried out using \texttt{R}, a widely used programming language for statistical calculations and visualizations.
%%
The interactive visualizations are realized using the \texttt{ggvis}\footnote{Developed by RStudio, Inc; \texttt{ggvis.rstudio.com}} package.
%%
The web-based approach allows for quick exchange with our collaborating epidemiological experts.
%%
They can use the technique without installing any software.
%%---------
\section{\uppercase{Results}}
\begin{figure*}[!t]
  \centering
  \includegraphics[width=1\textwidth]{figures/results}
  \caption{
%%
Decision tree quality plot of all classification results.
%%
The x-axis shows the number of decisions of the underlying model, the y-axis the classification error (see Section~\ref{sec:DecisionTrees}).
%%
The upper plot shows the results for all variables, either metrics expressed via their quantiles, or categorical.
%%
The lower plot displays the dummy variables derived from the original variables.
%%
Group affiliation of a data point is color coded:
%%
no group (a), subdivision into \emph{male} and \emph{female} subjects (b) and quartiles of \emph{Body Mass Index} (\emph{BMI}) (c).
%%
The number of subjects represented in a variable is denoted using the dot diameter.
%%
We only display variables with an error below $0.5$, results above this threshold are dismissed.
%%
The interactive plot (see supplemental material) has clickable data points, displaying the corresponding decision tree in a tool tip.
%%
Results of high interest are located close to the axis origins.
}
  \label{fig:results}
\end{figure*}
\noindent In this section, we show which non-image variables can be explained using the 9 image derived variables.
%%
We create subject groups to assess the influence of variables affecting the lumbar spine shape.
%%
The groups are: \emph{all subjects}; \emph{males} and \emph{females}; and a subdivision by \emph{Body Mass Index quantiles} ($BMI = \frac{m}{l^2}$ where $m$ is the \emph{body mass} in kilogram and $l$ is the \emph{body size} in meter), yielding the groups (17, 24.7] (24.7, 27.4] (27.4, 30.5] (30.5, 48].
%%
We plotted each group twice.
%%
The first plot shows all original variables, the second all categorical variables transformed into dichotomous \emph{dummy variables}.
%%
The shown mutual dependencies aim to amplify \emph{hypotheses generation}.
%%
Dedicated statistical analysis of these results solely focussed on epidemiological research is not the scope of this paper.
%%
The resulting plots can be seen in Figure~\ref{fig:results}.
%%
\\\\
\noindent \textbf{All Variables.}
%%
The majority of non-image variables cannot be automatically classified based on the image-variables.
%%
This is reflected in the large amount of variables classified with an error above $0.6$.
%%

None of the pain indicators can be reliably described using the image-based variables.
%%
The only variable reliably classified in this group is \emph{gender}, which can be described with an error of $0.31$ using 7 rules and incorporates only \emph{curvature-} and \emph{curvature angle}-related variables.
%%
The distinctness lies in the average difference in \emph{body size} between \emph{male} and \emph{female} subjects.
%%
\emph{Medication for high blood pressure} is classified for 1,058 subjects with an error of $0.47$ solely based on \emph{coronal mean curvature}.
%%
A high share of medicated subjects were correctly classified ($796$ of $1,058$).
%%
The majority of non-medicated subjects are false-positive classified ($262$ of $1,058$), yielding a poor quality of the classifier w.r.t. epidemiological research.
%%
The four \emph{body size} groups could be described with an error of $0.48$, but the decision tree comprises 71 rules and imposes overfitting.
%%
The dummy variable analysis yields a result similar to the \emph{blood pressure medication}.
%%
Variables, such as subject size 139-164 cm, between 64 and 90 years of \emph{age} or \emph{nutrition}-related variables are dominantly populated by one manifestation.
%%
The classifier neglects the other groups and yields an error below $0.5$.
%%
\\\\
\noindent \textbf{Gender Groups.}
%%
Classifications using groups divided by \emph{gender} do not produce satisfying results.
%%
Only \emph{hypothyroidism} could be described for male subjects with an error of $0.24$ for 110 subjects using the \emph{mean curvature} and \emph{curvature angle}.
%%
Since there are only 30 male subjects diagnosed with \emph{hypothyroidism}, the statistical power of the result is reduced.
%%
The dummy variable analysis showed that female subjects of 139-164~cm \emph{body size} could be discriminated using the \emph{mean curvature} and \emph{curvature angle}, with an error of $0.38$.
%%
\\\\
\noindent \textbf{Body Mass Index Groups.}
%%
\emph{Gender} could be classified for each \emph{BMI} group using \emph{mean curvature} and \emph{curvature angle}.
%%
The error varies between $0.31$ (\emph{BMI} of \emph{$30.5-48$}, 4 decision rules) to $0.39$ (\emph{BMI} of \emph{$24.7-27.4$}, 5 decision rules).
%%
The starting age of smoking could be described well with an error between $0.25$ and $0.32$ for all \emph{BMI} groups except for subjects with a \emph{BMI} of \emph{$30.5-48$}.
%%
The result is overfitted to the data due to tree sizes between $14$ and $16$.
%%
Some variables, such as \emph{body size}, can be described with an error of $0.3$ to $0.36$ using large decision trees with over 20 rules.
%%
Using mostly \emph{mean curvature} and \emph{curvature angle}, the \emph{leg pain level} can be described using $14$ rules with an error of $0.46$ for obese subjects (\emph{BMI} higher than $30$).
%%
This result also imposes overfitting.
%%
Subjects experiencing \emph{pain in the last seven days} can also be described for this group using the same variables with a tree consisting of $8$ rules and an error of $0.35$.
%%
Obese subjects are prone to \emph{back} and \emph{leg pain} due to a more stressed lumbar spine.
%%
The stress-induced spine deformation seems to directly influence the pain levels for these subjects.
%%
The dummy variable analysis shows many results using a decision tree with one rule based on \emph{mean curvature} or \emph{curvature angle} with an error between $0.35$ and $0.47$.
%%
\section{\uppercase{Summary \& Conclusion}}
\label{sec:Conclusion}
%% Contributions
\noindent We provide a method for comparative analysis of decision trees independent of the variable manifestation count using a novel \emph{decision tree quality plot}.
%%
We applied the method to gain insight into the predictive power of 9 image-derived variables for 134 non-image variables with focus on \emph{back pain}.
%%
The analysis was performed for subject groups of \emph{gender}, \emph{BMI} and \emph{body size} to assess their influence on the lumbar spine shape.

The methods presented herein may be applied to comprehensive epidemiological data sets to investigate mutual dependencies among variables and to generate hypotheses on potential associations and subgroups.
%%
These hypotheses, however, have to be substantiated by dedicated statistical analyses and replication in independent cohorts.
%%
\\\\
\noindent \textbf{Predictive Power of Image-Derived Variables.}
The presented results indicate that \emph{torsion}, \emph{curvature} and \emph{curvature angle} of the lumbar spine at the presented precision are not sufficient to describe lumbar back pain in the SHIP data set.
%%
Our method allows to assess their discriminative power, which is largely limited to separating male and female subjects, \emph{nutrition} variables, as well as different disease indicators.
%%
The C4.5 algorithm proved to be an effective tool for evaluating a set of derived metrics regarding their suitability to classify non-image variables.
%%
Over-fitting to the data indicated by complex decision trees has to be taken into account as well.
%%
The presented method only captures linear relationships between variables.
%%
To take more complex associations into account, methods such as regression analysis have to be incorporated.
\\\\
\noindent \textbf{Applicability.}
%%
Methods supporting \emph{hypotheses generation} based on image information are new to the application domain.
%%
They are an addition to the standard epidemiological workflow as they highlight new and possibly unknown relationships.
%%
Classification methods based on decision trees have proven to be useful for assessing the discriminative power of a variable set.
%%
Their ability to consider variable combinations makes them more powerful than correlation coefficients calculated for each variable.
%%
This advantage comes with a much more complex output, the results are more challenging to assess and to abstract.
%%
Our method to plot derived metrics and custom-tailored error measures proved to be effective.
%%
Huge result spaces could be navigated fast using our decision tree quality plot.
%%
Therefore, the method is applicable not only for deriving information based on image data, but on all \emph{potential target variables}.
%%
\\\\
\noindent \textbf{Future Work.}
We will focus on more precise models for extracting measures.
%%
Dented vertebrae indicate pathological deformation, and can be captured by segmenting the top and bottom point of each vertebra center.
%%
Spine canal thickness indicates signs of herniated disc disease and is also of interest.
%%
We aim to include the method into existing visual analytics methods designed for analyzing shape information for epidemiological data \cite{Klemm2014VIS}.
%%
\\\\
\noindent \textbf{Outlook.}
%%
Combining the power of statistical analyses, visual analytics and classification techniques is essential for analyzing increasingly complex heterogenous population data.
%%
These methods do not aim to replace the traditional epidemiological workflow, but rather complement the weak points of standard statistical methods.
%%
Our method provides a novel way to gain insight into these complex data sets and amplifies \emph{hypotheses generation}.
%%
\section*{\uppercase{Acknowledgements}}
\begin{small}
\noindent SHIP is part of the Community Medicine Research net of the University of Greifswald, Germany, which is funded by the Federal Ministry of Education and Research (grant no. 03ZIK012), the Ministry of Cultural Affairs as well as the Social Ministry of the Federal State of Mecklenburg-West Pomerania. Whole-body MR imaging was supported by a joint grant from Siemens Healthcare, Erlangen, Germany and the Federal State of Mecklenburg-Vorpommern. The University of Greifswald is a member of the ‘Centre of Knowledge Interchange’ program of the Siemens AG. This work was supported by the DFG Priority Program 1335: Scalable Visual Analytics. This work was supported by the federal state of Saxony-Anhalt under grant number 'I 60' within the Forschungscampus STIMULATE.
\end{small}

\bibliographystyle{style/apalike}
{\small
\bibliography{bibliography}}

\vfill
\end{document}